Optimal Fitting of Planar Curves to Prescribed Constraints

### 2013-08-08 idp/documentation/paper.tex
continued working on section about spiro splines

			\subsubsection{Common Design Tools for Bézier Splines}

@@

				The language used to specify Bézier splines in common design tools is lacking in expressiveness. Specifically, describing curves in terms of their curvature is difficult since curvature may only be specified indirectly through the magnitude of the velocity values (higher velocity resulting in lower curvature). Unfortunately, this is also very dependent on the point and velocity coefficients at the other end of the spline segment. There is also no way to directly specify linearly changing curvature. This makes specifying curvature very tedious, and in case the winding angle of the curve segment is large enough, actually impossible, forcing the designer to use additional nodes in order to get the desired result. Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created, or moved in order to change the overall curve. It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes. Another issue is that without software support for curvature continuity, using more nodes usually also introduces more points where curvature is not continuous. In cases where the curvature is an important property of the curve (for example the constant curvature of circle arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points, making it easy to make small mistakes.

@@

				Example 1 shows two Bézier splines. The left curve is a circle arc, exhibiting constant curvature, while the right one is not, going from low curvature to high curvature and back to low curvature. Since the difference can be hard to see, it's easy to make mistakes when trying to create circle arcs.

@@

				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_curve_design_tools}. The language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them being needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3). The fairness measure leads to beautifully smooth curves that are ususally sufficiently close to what a human would expect given the control points (criterion 4). The software is also very good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 5 well enough.

				The only real issue lies in the expressiveness of the language.

				% would in principle be well-suited for circle arcs and spirals and other shapes with linearly changing curvature, but curvature can only be controlled indirectly using points
				% tangent direction and curvature can only be controlled indirectly
				%   sometimes, many nodes need to be placed to convey the desired direction/curvature information
				%   specifying direction/curvature always comes with the unwanted side-effect of specifying additional points, possibly introducing unwanted bumpiness
				% sometimes, nice, smooth curves can be achieved using only few points
				%   it's not obvious where the points have to be specified such that one needs to specify as few as possible
				%   changing those 'minimal' curves is somewhat hampered by the fact that the purpose of each point may be non-obvious
				%     has bad locality since the points on the curve don't really specify points, but have an ulterior purpose
				% suffers from instability issues
				% does guarantee curvature continuity

>>>>

			\subsubsection{Common Design Tools for Bézier Splines}
			\label{section:common_design_tools_bézier_splines}

@@

				The language used to specify Bézier splines in common design tools is lacking in expressiveness. Specifically, describing curves in terms of their curvature is difficult since curvature may only be specified indirectly through the magnitude of the velocity values (higher velocity resulting in lower curvature). Unfortunately, this is also very dependent on the point and velocity coefficients at the other end of the spline segment. There is also no way to directly specify linearly changing curvature. This makes specifying curvature very tedious, and in case the winding angle of the curve segment is large enough, actually impossible, forcing the designer to use additional nodes in order to get the desired result. Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created or moved in order to change the overall curve. It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes. Another issue is that without software support for curvature continuity, using more nodes usually also introduces more points where curvature is not continuous. In cases where the curvature is an important property of the curve (for example the constant curvature of circle arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points.

@@

				Example 1 shows two Bézier splines. The left curve is a Bézier spline approximation of a circle arc, exhibiting nearly constant curvature, while the right one is not, going from low curvature to high curvature and back to low curvature. Since the difference can be hard to see, it's easy to make mistakes when trying to approximate circle arcs using Bézier splines.

@@

				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_curve_design_tools}. The language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them being needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3). The fairness measure leads to beautifully smooth curves that are ususally sufficiently close to what a human would expect given the control points (criterion 4). The software is also good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 5 well enough.

				The only real issue lies in the insufficient expressiveness of the language. Section \ref{section:common_design_tools_bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability. Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify direction nor curvature. These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature. This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to describing circle arcs, spirals and other shapes exhibiting constant or linearly changing curvature. Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties. Unfortunately, this is rarely practical since it both requires a lot of time to set up and doesn't behave very well when modified. These issues are somewhat alleviated by the guaranteed \(G^2\) continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved. That way, when indirectly specifying direction and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.
